Evolving Robot Gaits in Hardware: the HyperNEAT Generative Encoding Vs. Parameter Optimization

### 2011-06-10 paper_ecal/hyperNeatMethod.tex
Changes per three ECAL reviewers. Package for chairs started.

As with the parameterized methods, three runs of HyperNEAT
were performed. Runs lasted 20 generations and the population size was 9 (which allowed a bare minimum of diversity within and between NEAT species).  These numbers are small, but were necessarily
constrained given how much time it took to conduct evolution directly
on a real robot. The remaining parameters were identical to Clune et
al.~(2011).

>>>>

As with the parameterized methods, three runs of HyperNEAT were
performed. Runs lasted 20 generations with a population size of 9
organisms in 3 species, allowing a bare minimum of diversity
within and between NEAT species.  These numbers were
necessarily small given how much time it took to conduct
evolution directly on a real robot. The remaining parameters were
identical to Clune et al.~(2011).

### 2011-06-10 paper_ecal/learningMethods.tex
Changes per three ECAL reviewers. Package for chairs started.

We now describe the gait-generating algorithm classes.

@@

Intuitively, SineModel5 starts with 8 identical sine waves of
amplitude $\amp$ and period $\tau$, multiplies the waves for all outer
motors by $m_O$, multiplies the waves for all front motors by $m_F$,
and multiplies the waves for all right motors by $m_R$.  To obtain the
actual motor position commands, these waves are offset by
appropriate fixed constants ($C_O = 40$ for outer
motors, $C_I = 800$ for inner motors, and $C_C = 512$ for the center hip motor) so that the base position (when the sine
waves are at 0) is approximately a crouch (the position shown in
\figref{robot_whitebg.jpg}).  In this motion model, the ninth
(center) motor is kept at a neutral position.  Thus, the commanded
position for each motor (numbered in \figref{topdown2.jpg}), as a vector function of time, are produced by the following nine functions (listed in the order):

\[
\vec{g}(t) =
\left[ {\begin{array}{c@{ }l@{ }l}
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \           m_F            & + C_I \\ % 0
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \ m_O \cdot m_F            & + C_O \\ % 1
\amp \cdot \sin(2\pi t / \tau) & \                                  & + C_I \\ % 2
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \ m_O                      & + C_O \\ % 3
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \                     m_R  & + C_I \\ % 4
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \ m_O           \cdot m_R  & + C_O \\ % 5
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \           m_F \cdot m_R  & + C_I \\ % 6
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \ m_O \cdot m_F \cdot m_R  & + C_O \\ % 7
0                              & \                                  & + C_C \\ % 8
\end{array} } \right]
\]

@@

\emph{Nelder-Mead simplex method}: The Nelder-Mead simplex
  method \citep{nm} creates an initial simplex with $d+1$ vertices,
  where $d$ is the dimension of the parameter space. It then tests the
  fitness of each vertex, and, in general, it reflects the worst point
  over the centroid in an attempt to improve it.  Several additional
  rules are used to prevent cycles and local minima \citep{nm}.

>>>>

We now describe the classes of gait-generating algorithms.

@@

Intuitively, SineModel5 starts with 8 identical sine waves of
amplitude $\amp$ and period $\tau$, multiplies the waves for all outer
motors by $m_O$, multiplies the waves for all front motors by $m_F$,
and multiplies the waves for all right motors by $m_R$.  To obtain the
actual motor position commands, these waves are offset by
fixed constants ($C_O = 40$ for outer
motors, $C_I = 800$ for inner motors, and $C_C = 512$ for the center hip motor) so that the base position (when the sine
waves are at 0) is approximately a crouch (the position shown in
\figref{robot_whitebg.jpg}).
To keep the size of the model search space as
small as possible, we decided to keep the ninth (center) motor at a fixed neutral
position.
Thus, the commanded position for each motor as a vector function of
time is as follows (numbered as in \figref{topdown2.jpg}):

\[
\vec{g}(t) =
\left[ {\begin{array}{c@{ }c@{ }c@{ }l@{ }l}
\amp \cdot \sin(2\pi t / \tau) & \             & \ \cdot \ m_F & \            & + C_I \\ % 0
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \ m_O & \ \cdot \ m_F & \            & + C_O \\ % 1
\amp \cdot \sin(2\pi t / \tau) & \             & \             & \            & + C_I \\ % 2
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \ m_O & \             & \            & + C_O \\ % 3
\amp \cdot \sin(2\pi t / \tau) & \             & \             & \ \cdot \ m_R  & + C_I \\ % 4
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \ m_O & \             & \ \cdot \ m_R  & + C_O \\ % 5
\amp \cdot \sin(2\pi t / \tau) & \             & \ \cdot \ m_F & \ \cdot \ m_R  & + C_I \\ % 6
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \ m_O & \ \cdot \ m_F & \ \cdot \ m_R  & + C_O \\ % 7
0                              & \             & \             & \            & + C_C \\ % 8
\end{array} } \right]
\]

@@

\emph{Nelder-Mead simplex method}: The Nelder-Mead simplex
  method creates an initial simplex with $d+1$ vertices,
  for a $d$ dimensional parameter space. It then tests the
  fitness of each vertex, and, in general, it reflects the worst point
  over the simplex's centroid in an attempt to improve it.  Several additional
  rules are used to prevent cycles and local minima; see \cite{nm} for more information.
